---
title: "Multi-Scale Spatial Analysis for Nature-Based Urban Heat Island Mitigation"
subtitle: "ARFW0501 – Group D Project Report"
author:
  - name: "Artemi Kurski"
    affiliation: "MSc Geomatics, TU Delft"
  - name: "Niranjan Pradeep"
    affiliation: "MSc Geomatics, TU Delft"
  - name: "Darius Floroiu"
    affiliation: "MSc Geomatics, TU Delft"
  - name: "Henryk Gujda"
    affiliation: "MSc Geomatics, TU Delft"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
    include-in-header: components/styles.html
---


# Introduction

In the context of rapid urbanisation and climate change, it is increasingly important to understand how cities shape their local climate.

Comparing cities helps us see how differences in urban form influence temperature, environmental quality, and overall liveability. Cities differ in density, land use, building materials, and the amount of green space they contain. These factors strongly affect how heat is absorbed, stored, and released, which leads to noticeable temperature differences between urban areas. A key concept here is the **Urban Heat Island (UHI)** effect, where cities become warmer than their surrounding rural areas. This happens because materials like concrete and asphalt absorb and hold heat, vegetation is limited and cannot provide enough cooling through shade, and human activities such as transport, buildings, and industry release additional heat. Together, these factors increase local temperatures and can lead to higher heat stress, energy use, and air pollution.

Understanding UHI highlights the direct relationship between urban design and local climate. In this context, we aim to investigate how spatial data resolution influences UHI estimation and how urban blue-green infrastructure contributes to temperature mitigation.
In light of this we aim to explore the following research question:

```{=html}
{{< include components/research-question.html >}}
```

::: {.note-box}
**City Selection:** 
To answer the research question, we focus on two case study cities: Rotterdam and Guangzhou. These cities were selected for their similarities in coastal location and economic significance, while also exhibiting distinct urban forms and climates. `Rotterdam`, located in the Netherlands, is known for its modern architecture and extensive port facilities, while `Guangzhou`, in southern China, features a mix of traditional and contemporary urban development. By comparing these two cities, we can gain insights into how different urban designs and green infrastructure strategies impact UHI intensity across varying spatial resolutions.
:::
---

## Methods

Our methodology relies on a dual - scale approach to understand the urban climate. We analyzed both macro - level administrative boundaries (Dutch *buurten* and Chinese *jiedao*) and micro - level street networks. By sticking entirely to open-source data—primarily OpenStreetMap (OSM) and Landsat satellite imagery, we ensured that the pipeline built for Rotterdam could be directly transferred to Guangzhou without coordinate or data availability issues.

# Getting started

The first step of the analysis involves identifying the required datasets and defining the spatial boundaries of the study area.

For this purpose, **Google Earth Engine (GEE)** was used to access and process the main datasets. These datasets form the basis for the initial setup and help in selecting the relevant neighbourhoods for further analysis.


**1. Land Surface Temperature (LST):**
 LST was derived from Landsat 8 data in Google Earth Engine. Thermal bands were processed to obtain surface temperature, which was then used to study spatial heat variations across Rotterdam.

**2. NDVI:**
 Vegetation cover was computed from Landsat 8 spectral bands in Google Earth Engine using the Normalized Difference Vegetation Index (NDVI), providing a measure of urban greenness.

**3. Local Climate Zones (LCZ):**
 Urban form classification was obtained using the RUBCLIM LCZ dataset available in Google Earth Engine, which categorises urban areas based on built type and land cover characteristics.

**4. Administrative boundaries:**
 Municipal and neighbourhood boundaries for Rotterdam were taken from the `CBS Wijken en Buurten 2025 GeoPackage`, used to define the spatial extent of the study area.

**5. Street network:**
 The pedestrian street network was extracted from OpenStreetMap using the `osmnx` library, and used for analysing urban connectivity and spatial structure.


# Analysis Workflow
```{=html}
{{< include components/pipeline-widget.html >}}
```

## Typology construction

Instead of designing climate adaptations for thousands of individual streets or dozens of neighborhoods one by one, we grouped the urban fabric into distinct "typologies." These are essentially microclimate personas: clusters of areas that share the same environmental and spatial DNA. If we know what works for a specific typology in Rotterdam, we can test if that same solution works for its twin typology in Guangzhou.

We constructed these typologies across two distinct spatial scales:

**1. Data Acquisition and Indicator Processing**  
First, we established the thermal and physical baseline of the cities. Using Google Earth Engine, we pulled summer 2025 data from Landsat 8 (30m resolution) to calculate two key metrics:

*   **Vegetation (NDVI):** To measure the density of green infrastructure.
*   **Land Surface Temperature (LST):** To map the thermal landscape. To ensure accuracy across different satellite capture dates, we normalized the LST data by calculating the deviation from the spatial mean of each image, giving us a clean look at localized heat anomalies rather than day-to-day weather spikes. 

For urban morphology, we integrated the RUBCLIM Local Climate Zone (LCZ) dataset (100m resolution), which categorizes the physical build of the city into classes ranging from "Compact High-Rise" to "Sparsely Built."

**2. Pedestrian Accessibility (PST Engine)**  
To figure out how human-friendly these environments actually are, we couldn't just look at satellite pixels; we had to simulate how people move. Using OSM walking networks, we built a Python-based graph network to calculate Pedestrian Space Syntax (PST) accessibility. 

We identified "cooling hubs" (neighborhoods with an NDVI above the city median) and mapped their center points. The algorithm then simulated human navigation, counting the number of topological turns required to reach these cooling zones from any point on the street network. Paths requiring fewer than three turns received higher accessibility scores, mathematically penalizing areas that are cut off from blue-green infrastructure. 

**3. Macro-Scale Clustering (Neighborhoods)**  
With our variables set (Mean LST, Mean NDVI, LCZ class, and PST score), we aggregated the data up to the neighborhood level using zonal statistics. Because our data was mixed (blending continuous numbers like temperature with categorical tags like LCZ codes), standard clustering methods wouldn't work. 

Instead, we ran the data through R using a Gower distance matrix, which is specifically designed to handle mixed data types. We then applied Partitioning Around Medoids (PAM) clustering to group the neighborhoods into four distinct microclimate profiles.

**4. Micro-Scale Clustering (Street Segments)**  
Finally, we brought the analysis down to the human scale to find the most vulnerable individual streets. Using Python and Geopandas, we slightly buffered the OSM street linestrings and extracted the exact LST and NDVI pixel values sitting directly under the pavement, while sampling the LCZ classification at the start point of each street. 

After dealing with missing values and standardizing the numeric data, we used an elbow plot to determine the optimal number of clusters and ran a MiniBatch K-Means clustering algorithm. To verify that our math matched physical reality, the pipeline randomly sampled 50 streets from our most extreme clusters and exported their coordinates to a CSV, allowing us to drop into Google Street View and qualitatively inspect the physical characteristics of the hottest streets in the city.

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

# Read the layer from the GeoPackage
invisible(capture.output(
  roads_data <- st_read("data/rotterdam/rotterdam_roads_clustered.gpkg",
                        layer = "rotterdam_roads_data")
))

# Randomly select 1000 rows
set.seed(42)  # remove or change this if you don't want reproducible sampling
roads_sample <- roads_data %>% slice_sample(n = 1000)

# Colors for clusters 0-5
cluster_colors <- c(
  "0" = "#4C72B0",
  "1" = "#65C3A1",
  "2" = "#FC9964",
  "3" = "#869DC5",
  "4" = "#E597C0",
  "5" = "#D9C04C"
)

cluster3d_scaled <- plot_ly(
  x = roads_sample$lst,
  y = roads_sample$ndvi,
  z = roads_sample$Bww1kM,
  type = "scatter3d",
  mode = "markers",
  color = as.factor(roads_sample$cluster),
  colors = cluster_colors,
  marker = list(size = 3)
) %>%
  layout(
    title = "Street clusters in Rotterdam",
    scene = list(
      xaxis = list(title = "LST"),
      yaxis = list(title = "NDVI"),
      zaxis = list(title = "Betweenness")
    )
  )
cluster3d_scaled
```


## Small-scale output: Rotterdam

After completing step 1 and 2, the following results were obtained, which formed the basis for further analysis.

```{=html}
{{< include components/rotterdam-maps.html >}}
```


From the analysis, three neighbourhoods stand out as showing extreme variation relative to the rest of the city: **Bloemhof, Afrikaanderwijk, and Hillesluis**.
```{=html}
{{< include components/rotterdam-cluster-maps.html >}}
```






## Small-scale output: Guangzhou

```{=html}
{{< include components/guangzhou-maps.html >}}
```


```{=html}
{{< include components/yuexiu-district-china.html >}}
```


# Results

## Rotterdam: Street-view and Nature based solutions

```{=html}
{{< include components/street-view-rotterdam.html >}}
```

## Gunagzhou: Street-view and Nature based solutions

```{=html}
{{< include components/yuexiu-street-view.html >}}
```


```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

# Read the layer from the GeoPackage
invisible(capture.output(
  gzhou_roads_data <- st_read("data/guangzhou/guangzhou_roads_clustered.gpkg",
                        layer = "guangzhou_roads_data")
))

# Randomly select 1000 rows
set.seed(42)  # remove or change this if you don't want reproducible sampling
gzhou_roads_sample <- gzhou_roads_data %>% slice_sample(n = 1000)

# Colors for clusters 0-5
cluster_colors <- c(
  "0" = "#4C72B0",
  "1" = "#65C3A1",
  "2" = "#FC9964",
  "3" = "#869DC5",
  "4" = "#E597C0",
  "5" = "#D9C04C"
)

cluster3d_scaled <- plot_ly(
  x = gzhou_roads_sample$lst,
  y = gzhou_roads_sample$ndvi,
  z = gzhou_roads_sample$Bww1kM,
  type = "scatter3d",
  mode = "markers",
  color = as.factor(gzhou_roads_sample$cluster),
  colors = cluster_colors,
  marker = list(size = 3)
) %>%
  layout(
    title = "Street clusters in Guangzhou",
    scene = list(
      xaxis = list(title = "LST"),
      yaxis = list(title = "NDVI"),
      zaxis = list(title = "Betweenness")
    )
  )
cluster3d_scaled
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

# Convert a hex color to an rgba() string with a given alpha (reused from earlier)
hex_to_rgba <- function(hex, alpha) {
  rgb_vals <- col2rgb(hex)
  sprintf("rgba(%d,%d,%d,%.2f)", rgb_vals[1], rgb_vals[2], rgb_vals[3], alpha)
}

# Draws a violin SHAPE from a precomputed density curve (n points) instead of
# letting plotly's type="violin" embed every raw row to recompute the KDE
# client-side. This is what actually kills the file-size blow-up: regardless
# of whether `x` has 500 or 350,000 values, density() always returns `n`
# points, so the embedded JSON payload stays tiny and constant-size.
add_violin_from_density <- function(p, x, center, color, name,
                                     half_width = 0.35, n = 256,
                                     line_opacity = 0.8, fill_opacity = 0.4,
                                     showlegend = TRUE) {
  x <- x[is.finite(x)]
  d <- density(x, n = n)                      # ~256 (x,y) pairs, not 1 per row

  scale <- half_width / max(d$y)               # normalize to a fixed max width
  half  <- d$y * scale

  # closed polygon: left edge going up the value range, right edge coming back down
  poly_x <- c(center - half, rev(center + half))
  poly_y <- c(d$x,           rev(d$x))

  p %>% add_trace(
    type = "scatter", mode = "lines",
    x = poly_x, y = poly_y,
    fill = "toself",
    fillcolor = hex_to_rgba(color, fill_opacity),
    line = list(color = hex_to_rgba(color, line_opacity)),
    name = name, legendgroup = name, showlegend = showlegend,
    hoverinfo = "skip"
  )
}

# Box + mean line drawn from a 5-number summary, not the raw vector. plotly's
# box trace supports precomputed q1/median/q3/lowerfence/upperfence/mean
# directly, so this needs only ~6 numbers per group.
add_box_summary <- function(p, x, center, color, name, box_width = 0.12) {
  x  <- x[is.finite(x)]
  qs <- quantile(x, probs = c(0, 0.25, 0.5, 0.75, 1), names = FALSE)

  p %>% add_trace(
    type = "box",
    x0 = center, width = box_width,
    q1 = list(qs[2]), median = list(qs[3]), q3 = list(qs[4]),
    lowerfence = list(qs[1]), upperfence = list(qs[5]),
    mean = list(mean(x)), boxmean = TRUE,
    fillcolor = "white",
    line = list(color = color, width = 1),
    name = name, legendgroup = name, showlegend = FALSE,
    hoverinfo = "skip"
  )
}

# Violin plots for LST, NDVI, and betweenness (Bww1kM), one subplot per
# variable, one violin per city within each
# data_list: named list, e.g. list(Rotterdam = roads_data, Guangzhou = gzhou_roads_data)
plot_lst_ndvi_violin <- function(data_list, colors = c("#4C72B0", "#FC9964"),
                                  title = "Distribution of LST, NDVI, and Betweenness by city") {

  # Build one violin subplot for a given variable (e.g. "lst", "ndvi", "Bww1kM")
  build_panel <- function(var, y_title, show_legend) {
    p <- plot_ly()
    for (i in seq_along(data_list)) {
      city_name <- names(data_list)[i]
      x <- data_list[[city_name]] %>% st_drop_geometry() %>% pull(.data[[var]])

      p <- p %>%
        add_violin_from_density(x, center = i, color = colors[i], name = city_name,
                                 showlegend = show_legend) %>%
        add_box_summary(x, center = i, color = colors[i], name = city_name)
    }
    p %>% layout(
      yaxis = list(title = y_title),
      xaxis = list(title = "", tickvals = seq_along(data_list),
                   ticktext = names(data_list), range = c(0.5, length(data_list) + 0.5))
    )
  }

  p_lst  <- build_panel("lst",    "LST",         show_legend = TRUE)
  p_ndvi <- build_panel("ndvi",   "NDVI",        show_legend = FALSE)
  p_bw   <- build_panel("Bww1kM", "Betweenness", show_legend = FALSE)

  subplot(p_lst, p_ndvi, p_bw, nrows = 1, titleY = TRUE, margin = 0.06) %>%
    layout(
      title = title,
      legend = list(title = list(text = "City"))
    )
}

plot_lst_ndvi_violin(list(Rotterdam = roads_data, Guangzhou = gzhou_roads_data))
```

```{r message=FALSE, warning=FALSE}
# LST/NDVI correlation for Gzhou

library(plotly)

# Correlation coefficient
correlation <- cor(gzhou_roads_sample$lst, gzhou_roads_sample$ndvi, use = "complete.obs")

# Linear model for the trend line
lm_fit <- lm(ndvi ~ lst, data = gzhou_roads_sample)
lst_range <- seq(min(gzhou_roads_sample$lst, na.rm = TRUE),
                  max(gzhou_roads_sample$lst, na.rm = TRUE),
                  length.out = 100)
trend_pred <- predict(lm_fit, newdata = data.frame(lst = lst_range))

scatter_lst_ndvi <- plot_ly(
  data = gzhou_roads_sample,
  x = ~lst,
  y = ~ndvi,
  type = "scatter",
  mode = "markers",
  color = ~as.factor(cluster),
  colors = cluster_colors,
  marker = list(size = 6, opacity = 0.7),
  name = ~as.factor(cluster)
) %>%
  add_trace(
    x = lst_range,
    y = trend_pred,
    type = "scatter",
    mode = "lines",
    line = list(color = "#000000", width = 2, dash = "dash"),
    name = "Trend line",
    inherit = FALSE   # <-- prevents it from inheriting the marker settings above
  ) %>%
  layout(
    title = paste0("Guangzhou: LST vs NDVI (r = ", round(correlation, 3), ")"),
    xaxis = list(title = "LST"),
    yaxis = list(title = "NDVI")
  )

scatter_lst_ndvi
```

```{r message=FALSE, warning=FALSE}
# LST/NDVI correlation for Rdam
library(plotly)

# Correlation coefficient
correlation <- cor(roads_sample$lst, roads_sample$ndvi, use = "complete.obs")

# Linear model for the trend line
lm_fit <- lm(ndvi ~ lst, data = roads_sample)
lst_range <- seq(min(roads_sample$lst, na.rm = TRUE),
                  max(roads_sample$lst, na.rm = TRUE),
                  length.out = 100)
trend_pred <- predict(lm_fit, newdata = data.frame(lst = lst_range))

scatter_lst_ndvi <- plot_ly(
  data = roads_sample,
  x = ~lst,
  y = ~ndvi,
  type = "scatter",
  mode = "markers",
  color = ~as.factor(cluster),
  colors = cluster_colors,
  marker = list(size = 6, opacity = 0.7),
  name = ~as.factor(cluster)
) %>%
  add_trace(
    x = lst_range,
    y = trend_pred,
    type = "scatter",
    mode = "lines",
    line = list(color = "#000000", width = 2, dash = "dash"),
    name = "Trend line",
    inherit = FALSE   # <-- prevents it from inheriting the marker settings above
  ) %>%
  layout(
    title = paste0("Rotterdam: LST vs NDVI (r = ", round(correlation, 3), ")"),
    xaxis = list(title = "LST"),
    yaxis = list(title = "NDVI")
  )

scatter_lst_ndvi
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

lcz_labels <- c(
  "0"  = "No data",
  "1"  = "Compact highrise",
  "2"  = "Compact midrise",
  "3"  = "Compact lowrise",
  "4"  = "Open highrise",
  "5"  = "Open midrise",
  "6"  = "Open lowrise",
  "7"  = "Lightweight lowrise",
  "8"  = "Large lowrise",
  "9"  = "Sparsely built",
  "10" = "Heavy industry",
  "11" = "Dense trees (LCZ A)",
  "12" = "Scattered trees (LCZ B)",
  "13" = "Bush, scrub (LCZ C)",
  "14" = "Low plants (LCZ D)",
  "15" = "Bare rock/paved (LCZ E)",
  "16" = "Bare soil/sand (LCZ F)",
  "17" = "Water (LCZ G)"
)
lcz_short_labels <- c("0" = "no data", setNames(paste("LCZ", 1:17), as.character(1:17)))

# % of a city's TOTAL road length per LCZ (no cluster grouping)
city_lcz_pct <- function(roads_data, all_lcz = names(lcz_labels)) {
  roads_data$seg_length <- as.numeric(st_length(roads_data))

  df <- roads_data %>%
    st_drop_geometry() %>%
    mutate(lcz = ifelse(is.na(lcz), "0", as.character(lcz))) %>%
    group_by(lcz) %>%
    summarise(total_length = sum(seg_length, na.rm = TRUE), .groups = "drop") %>%
    mutate(pct = 100 * total_length / sum(total_length))

  tibble(lcz = all_lcz) %>%
    left_join(df, by = "lcz") %>%
    mutate(
      pct = ifelse(is.na(pct), 0, pct),
      lcz_full = lcz_labels[lcz],
      lcz_short = lcz_short_labels[lcz]
    ) %>%
    arrange(match(lcz, all_lcz))
}

# Convert a hex color (e.g. "#4C72B0") to an rgba() string with a given alpha
hex_to_rgba <- function(hex, alpha) {
  rgb_vals <- col2rgb(hex)
  sprintf("rgba(%d,%d,%d,%.2f)", rgb_vals[1], rgb_vals[2], rgb_vals[3], alpha)
}

plot_city_lcz_radar <- function(data_list, colors = c("#4C72B0", "#FC9964"),
                                 line_opacity = 0.6, fill_opacity = 0.1,
                                 title = "LCZ composition of total road network length",
                                 title_pad = 16) {
  fig <- plot_ly(type = 'scatterpolar', fill = 'toself')

  for (i in seq_along(data_list)) {
    city_name <- names(data_list)[i]
    pct_df <- city_lcz_pct(data_list[[i]])

    theta_vals <- c(pct_df$lcz_short, pct_df$lcz_short[1])
    r_vals <- c(pct_df$pct, pct_df$pct[1])
    hover_labels <- c(pct_df$lcz_full, pct_df$lcz_full[1])

    fig <- fig %>%
      add_trace(
        r = r_vals,
        theta = theta_vals,
        name = city_name,
        text = hover_labels,
        hovertemplate = paste0("<b>", city_name, "</b><br>%{text}: %{r:.1f}%<extra></extra>"),
        line = list(color = hex_to_rgba(colors[i], line_opacity)),
        marker = list(color = hex_to_rgba(colors[i], line_opacity)),
        fillcolor = hex_to_rgba(colors[i], fill_opacity)
        # no trace-level `opacity` — line/marker and fill now control their own alpha independently
      )
  }

  fig %>%
    layout(
      polar = list(
        domain = list(y = c(0, 0.88)),
        radialaxis = list(visible = TRUE, ticksuffix = "%")
      ),
      title = list(text = title, y = 0.97, yanchor = "top"),
      margin = list(t = title_pad)
    )
}

plot_city_lcz_radar(list(Rotterdam = roads_data, Guangzhou = gzhou_roads_data))
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

lcz_labels <- c(
  "0"  = "No data",
  "1"  = "Compact highrise",
  "2"  = "Compact midrise",
  "3"  = "Compact lowrise",
  "4"  = "Open highrise",
  "5"  = "Open midrise",
  "6"  = "Open lowrise",
  "7"  = "Lightweight lowrise",
  "8"  = "Large lowrise",
  "9"  = "Sparsely built",
  "10" = "Heavy industry",
  "11" = "Dense trees (LCZ A)",
  "12" = "Scattered trees (LCZ B)",
  "13" = "Bush, scrub (LCZ C)",
  "14" = "Low plants (LCZ D)",
  "15" = "Bare rock/paved (LCZ E)",
  "16" = "Bare soil/sand (LCZ F)",
  "17" = "Water (LCZ G)"
)

lcz_short_labels <- c("0" = "no data", setNames(paste("LCZ", 1:17), as.character(1:17)))

cluster_lcz_pct <- function(roads_data, all_lcz = names(lcz_labels)) {
  roads_data$seg_length <- as.numeric(st_length(roads_data))

  df <- roads_data %>%
    st_drop_geometry() %>%
    mutate(
      cluster = as.character(cluster),
      lcz = ifelse(is.na(lcz), "0", as.character(lcz))
    ) %>%
    group_by(cluster, lcz) %>%
    summarise(total_length = sum(seg_length, na.rm = TRUE), .groups = "drop") %>%
    group_by(cluster) %>%
    mutate(pct = 100 * total_length / sum(total_length)) %>%
    ungroup()

  clusters <- sort(unique(df$cluster))
  full_grid <- expand.grid(cluster = clusters, lcz = all_lcz, stringsAsFactors = FALSE)

  full_grid %>%
    left_join(df, by = c("cluster", "lcz")) %>%
    mutate(
      pct = ifelse(is.na(pct), 0, pct),
      lcz_full = lcz_labels[lcz],
      lcz_short = lcz_short_labels[lcz]
    ) %>%
    arrange(cluster, match(lcz, all_lcz))
}

# Faceted radar with explicit pixel sizing so panels stay square and fill their allotted space
plot_cluster_lcz_radar_facets <- function(roads_data, colors = cluster_colors,
                                           ncols = 2,
                                           panel_size = 280,   # px — each polar panel is panel_size x panel_size
                                           title_band = 40,    # px reserved above each panel for its title
                                           gap = 100,            # px gap between panels
                                           main_title_margin = 60,  # px reserved at top for the overall title
                                           side_margin = 48,
                                           bottom_margin = 20,
                                           title = "LCZ composition by cluster") {
  pct_df <- cluster_lcz_pct(roads_data)
  clusters <- sort(unique(pct_df$cluster))
  n <- length(clusters)
  nrows <- ceiling(n / ncols)

  r_max <- ceiling(max(pct_df$pct, na.rm = TRUE) / 5) * 5

  # Plot-area (inside margins) dimensions, built up from panel_size so squares stay square
  plot_area_width  <- ncols * panel_size + (ncols - 1) * gap
  plot_area_height <- nrows * (panel_size + title_band) + (nrows - 1) * gap

  fig_width  <- plot_area_width + 2 * side_margin
  fig_height <- plot_area_height + main_title_margin + bottom_margin

  fig <- plot_ly(type = 'scatterpolar', fill = 'toself')
  polar_layouts <- list()
  annotations <- list()

  for (i in seq_along(clusters)) {
    cl <- clusters[i]
    row_idx <- (i - 1) %/% ncols
    col_idx <- (i - 1) %% ncols

    x0_px <- col_idx * (panel_size + gap)
    x1_px <- x0_px + panel_size

    block_top_px    <- row_idx * (panel_size + title_band + gap)
    polar_top_px    <- block_top_px + title_band
    polar_bottom_px <- polar_top_px + panel_size

    # Fractions are relative to the plot area (same space "paper" coords use), not the full figure
    x0 <- x0_px / plot_area_width
    x1 <- x1_px / plot_area_width
    y1 <- 1 - polar_top_px / plot_area_height
    y0 <- 1 - polar_bottom_px / plot_area_height
    title_y <- 1 - block_top_px / plot_area_height

    polar_name <- if (i == 1) "polar" else paste0("polar", i)

    sub <- pct_df %>% filter(cluster == cl)
    theta_vals <- c(sub$lcz_short, sub$lcz_short[1])
    r_vals <- c(sub$pct, sub$pct[1])
    hover_labels <- c(sub$lcz_full, sub$lcz_full[1])

    fig <- fig %>%
      add_trace(
        r = r_vals,
        theta = theta_vals,
        subplot = polar_name,
        name = paste("Cluster", cl),
        text = hover_labels,
        hovertemplate = paste0("<b>Cluster ", cl, "</b><br>%{text}: %{r:.1f}%<extra></extra>"),
        line = list(color = colors[[cl]]),
        marker = list(color = colors[[cl]]),
        opacity = 0.7,
        showlegend = FALSE
      )

    polar_layouts[[polar_name]] <- list(
      domain = list(x = c(x0, x1), y = c(y0, y1)),
      radialaxis = list(visible = TRUE, range = c(0, r_max), ticksuffix = "%", tickfont = list(size = 8)),
      angularaxis = list(tickfont = list(size = 8))
    )

    annotations[[i]] <- list(
      text = paste("Cluster", cl),
      x = (x0 + x1) / 2, y = title_y,
      xanchor = "center", yanchor = "top",
      xref = "paper", yref = "paper",
      showarrow = FALSE,
      font = list(size = 13)
    )
  }

  layout_args <- c(
    list(fig),
    polar_layouts,
    list(
      title = title,
      annotations = annotations,
      width = fig_width,
      height = fig_height,
      margin = list(t = main_title_margin, b = bottom_margin, l = side_margin, r = side_margin)
    )
  )

  do.call(layout, layout_args)
}
```

```{r message=FALSE, warning=FALSE}
plot_cluster_lcz_radar_facets(roads_data, title = "Rotterdam: LCZ composition by cluster")
```

```{r message=FALSE, warning=FALSE}
plot_cluster_lcz_radar_facets(gzhou_roads_data, title = "Guangzhou: LCZ composition by cluster")
```

# Discussion
## Interpretation of the findings

The results demonstrate that urban heat island patterns can be identified effectively through a combination of remotely sensed environmental indicators and street-level spatial analysis. At the neighbourhood scale, the analysis of Land Surface Temperature (LST), Normalized Difference Vegetation Index (NDVI), and Local Climate Zones (LCZ) successfully identified broad spatial patterns of urban heat. Industrial districts and densely built urban areas generally exhibited higher surface temperatures and lower vegetation cover, while greener neighbourhoods consistently showed lower thermal intensity. These findings are consistent with established urban climate research, which identifies vegetation as an important contributor to surface cooling through shading and evapotranspiration.

However, one of the main findings of this study was that neighbourhood-level averages alone are insufficient for identifying appropriate intervention locations. While they effectively highlight broad hotspot areas, they conceal considerable thermal variability within individual neighbourhoods. By transferring the analysis to the street scale, it became possible to identify individual street segments experiencing elevated thermal exposure, providing a much more actionable basis for climate adaptation planning.

The integration of environmental indicators with street-level accessibility and urban morphology further demonstrated that high temperatures alone do not necessarily determine intervention priorities. Streets with similar thermal conditions often differed substantially in their surrounding built environment, vegetation availability, and spatial accessibility. Consequently, different hotspot environments require different nature-based solutions. Dense residential streets with limited available space were found to be more suitable for vertical greening solutions such as green façades and living walls, whereas wider streets offered opportunities for street tree planting. This highlights the importance of considering local urban context alongside thermal conditions when designing climate adaptation strategies.

Finally, the comparison between Rotterdam and Guangzhou suggests that, despite their contrasting climates and urban forms, the proposed analytical workflow is transferable between cities. Although the specific interventions should always reflect local urban conditions, the combination of satellite-derived environmental indicators and street-level analysis provides a flexible framework for identifying thermally vulnerable areas and supporting evidence-based nature-based solution planning.

## Multi-scale spatial analysis

One of the most significant findings of this study is the importance of spatial scale when analysing urban heat islands and identifying suitable intervention locations. At the neighbourhood scale, the analysis successfully identified broad thermal patterns and highlighted districts characterised by elevated land surface temperatures and limited vegetation cover. This level of aggregation provides a valuable overview of the spatial distribution of heat across the city and is well suited for identifying priority areas for further investigation.

However, neighbourhood averages conceal substantial variability within individual urban blocks. When the analysis was transferred to the street scale, considerably more detailed spatial patterns emerged. Rather than entire neighbourhoods exhibiting uniformly high temperatures, elevated thermal exposure was concentrated along specific streets and street segments. This demonstrates that neighbourhood-scale analyses alone are insufficient for guiding targeted climate adaptation measures, as interventions are ultimately implemented within the public realm at the scale of individual streets. From a planning perspective, neighbourhood-level analysis is valuable for prioritising investment, whereas street-level analysis provides the spatial precision required to determine where interventions should actually be implemented.

The comparison between Rotterdam and Guangzhou further illustrates how multi-scale analysis reveals differences in urban morphology that would not be apparent from city-wide statistics alone. While the same analytical workflow was successfully transferred to both cities, the resulting hotspot environments differed considerably.

In Rotterdam, the highest temperatures were predominantly observed within dense residential neighbourhoods, particularly along narrow side streets characterised by limited vegetation and enclosed urban form. Large commercial areas and major boulevards generally exhibited lower thermal intensity than expected, likely reflecting the extensive investment in blue-green infrastructure and mature street trees already present throughout the city. This suggests that previous investments in Rotterdam's blue-green infrastructure may already be mitigating heat accumulation along major transport corridors, shifting thermal vulnerability towards smaller residential streets where greening opportunities remain more limited.

By contrast, the Guangzhou case study revealed a different urban structure. Thermal hotspots were frequently associated with large transport corridors, major intersections and expansive road infrastructure surrounding residential developments. These environments are characterised by large impervious surfaces and considerably wider streets than those typically found in Rotterdam. Consequently, although the analytical framework remained transferable, the resulting intervention priorities differed substantially between the two cities. Unlike Rotterdam, where interventions primarily target dense residential streets, Guangzhou offers opportunities for larger-scale linear green infrastructure along major transport corridors. The same methodology therefore produced different adaptation strategies despite relying on identical analytical procedures.

## Nature-based solutions

The proposed nature-based solutions were selected according to the physical characteristics of each hotspot environment rather than thermal conditions alone. This study demonstrates that different urban forms require different approaches to urban cooling.

In Rotterdam, many of the identified hotspot streets were located within dense residential neighbourhoods where available public space is limited. Existing major boulevards and principal transport corridors already incorporate substantial tree cover and green infrastructure, reducing their relative thermal vulnerability. As a result, interventions focused on smaller residential streets where additional cooling potential remains. Green façades and living walls were proposed for locations with limited planting space, allowing vegetation to be introduced without significantly altering the existing street layout. Where sufficient pavement width existed, street trees and linear planting schemes were recommended to increase shading, improve pedestrian thermal comfort and enhance evapotranspiration.

The Guangzhou case study presented a contrasting urban context. Here, hotspot areas were frequently associated with large arterial roads, highway interchanges and expansive paved surfaces adjacent to residential districts. These environments provide considerably more physical space for intervention but also generate greater heat through extensive impervious cover. Consequently, larger-scale greening strategies may be more appropriate, including continuous roadside tree planting, vegetated buffers and extensive green walls alongside major transport corridors. In addition to reducing surface temperatures through shading and evapotranspiration, such interventions may also improve acoustic conditions by providing a vegetated barrier between traffic infrastructure and surrounding residential areas.

These findings emphasise that effective nature-based solutions cannot be selected solely on the basis of temperature. Instead, they should be matched to local urban morphology, available space and existing green infrastructure. While the specific interventions differed between Rotterdam and Guangzhou, the analytical workflow consistently supported the identification of locations where context-sensitive climate adaptation measures are likely to provide the greatest benefit.

## Limitations

A key limitation of this study is the reliance on Land Surface Temperature rather than direct measurements of pedestrian thermal comfort. Although LST provides valuable information regarding heat accumulation, human thermal perception is additionally influenced by air temperature, humidity, wind speed, solar radiation and shading.
Furthermore, the datasets used in this study were acquired at different spatial resolutions. Landsat imagery was available at a resolution of 30 m, whereas Local Climate Zones were represented at 100 m resolution. Although these datasets could be integrated successfully, aggregating them to street level inevitably introduces spatial uncertainty. Finally, the analysis represents thermal conditions during a limited observation period and therefore does not capture seasonal or diurnal variability. Future work could incorporate additional climatic variables, perform seasonal analyses and evaluate the cooling performance of proposed nature-based solutions through microclimate simulations.

# Conclusion

This study demonstrates that combining neighbourhood-scale and street-scale spatial analysis provides an effective framework for identifying urban heat hotspots and selecting context-sensitive nature-based solutions.
Neighbourhood-scale analysis successfully identified broad thermal patterns and priority areas for further investigation. However, transferring the analysis to the street scale revealed considerably greater spatial detail, allowing individual hotspot corridors and representative urban environments to be identified. This finer spatial resolution proved essential for selecting interventions that reflect the physical characteristics of each location rather than relying solely on aggregated neighbourhood statistics.
The comparison between Rotterdam and Guangzhou further demonstrated that, although urban morphology strongly influences the type of intervention required, the overall analytical workflow remains transferable across cities with different climatic and spatial characteristics. While Rotterdam's hotspot areas were predominantly located within dense residential neighbourhoods, Guangzhou's hotspots were largely associated with large transport corridors and major road infrastructure, leading to different but equally context-specific nature-based solution strategies.
Overall, the study highlights that broad-scale analysis identifies where urban heat occurs, while street-scale analysis identifies where meaningful interventions can be implemented. By combining open spatial datasets with multi-scale analysis, the proposed workflow provides a practical and transferable framework for supporting evidence-based urban climate adaptation and nature-based planning.
