---
title: "Multi-Scale Spatial Analysis for Nature-Based Urban Heat Island Mitigation"
subtitle: "ARFW0501 – Group D Project Report"
author:
  - name: "Artemi Kurski"
    affiliation: "MSc Geomatics, TU Delft"
  - name: "Niranjan Pradeep"
    affiliation: "MSc Geomatics, TU Delft"
  - name: "Darius Floroiu"
    affiliation: "MSc Geomatics, TU Delft"
  - name: "Henryk Gujda"
    affiliation: "MSc Geomatics, TU Delft"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
    include-in-header: components/styles.html
---


# Introduction

In the context of rapid urbanisation and climate change, it is increasingly important to understand how cities shape their local climate.

Comparing cities helps us see how differences in urban form influence temperature, environmental quality, and overall liveability. Cities differ in density, land use, building materials, and the amount of green space they contain. These factors strongly affect how heat is absorbed, stored, and released, which leads to noticeable temperature differences between urban areas. A key concept here is the **Urban Heat Island (UHI)** effect, where cities become warmer than their surrounding rural areas. This happens because materials like concrete and asphalt absorb and hold heat, vegetation is limited and cannot provide enough cooling through shade, and human activities such as transport, buildings, and industry release additional heat. Together, these factors increase local temperatures and can lead to higher heat stress, energy use, and air pollution.

Understanding UHI highlights the direct relationship between urban design and local climate. In this context, we aim to investigate how spatial data resolution influences UHI estimation and how urban blue-green infrastructure contributes to temperature mitigation.
In light of this we aim to explore the following research question:

```{=html}
{{< include components/research-question.html >}}
```

::: {.note-box}
**City Selection:** 
To answer the research question, we focus on two case study cities: Rotterdam and Guangzhou. These cities were selected for their similarities in coastal location and economic significance, while also exhibiting distinct urban forms and climates. `Rotterdam`, located in the Netherlands, is known for its modern architecture and extensive port facilities, while `Guangzhou`, in southern China, features a mix of traditional and contemporary urban development. By comparing these two cities, we can gain insights into how different urban designs and green infrastructure strategies impact UHI intensity across varying spatial resolutions.
:::
---

## Methods

Our methodology relies on a dual - scale approach to understand the urban climate. We analyzed both macro - level administrative boundaries (Dutch *buurten* and Chinese *jiedao*) and micro - level street networks. By sticking entirely to open-source data—primarily OpenStreetMap (OSM) and Landsat satellite imagery, we ensured that the pipeline built for Rotterdam could be directly transferred to Guangzhou without coordinate or data availability issues.

# Getting started

The first step of the analysis involves identifying the required datasets and defining the spatial boundaries of the study area.

For this purpose, **Google Earth Engine (GEE)** was used to access and process the main datasets. These datasets form the basis for the initial setup and help in selecting the relevant neighbourhoods for further analysis.


**1. Land Surface Temperature (LST):**
 LST was derived from Landsat 8 data in Google Earth Engine. Thermal bands were processed to obtain surface temperature, which was then used to study spatial heat variations across Rotterdam.

**2. NDVI:**
 Vegetation cover was computed from Landsat 8 spectral bands in Google Earth Engine using the Normalized Difference Vegetation Index (NDVI), providing a measure of urban greenness.

**3. Local Climate Zones (LCZ):**
 Urban form classification was obtained using the RUBCLIM LCZ dataset available in Google Earth Engine, which categorises urban areas based on built type and land cover characteristics.

**4. Administrative boundaries:**
 Municipal and neighbourhood boundaries for Rotterdam were taken from the `CBS Wijken en Buurten 2025 GeoPackage`, used to define the spatial extent of the study area.

**5. Street network:**
 The pedestrian street network was extracted from OpenStreetMap using the `osmnx` library, and used for analysing urban connectivity and spatial structure.


# Analysis Workflow
```{=html}
{{< include components/pipeline-widget.html >}}
```

## Typology construction

Instead of designing climate adaptations for thousands of individual streets or dozens of neighborhoods one by one, we grouped the urban fabric into distinct "typologies." These are essentially microclimate personas: clusters of areas that share the same environmental and spatial DNA. If we know what works for a specific typology in Rotterdam, we can test if that same solution works for its twin typology in Guangzhou.

We constructed these typologies across two distinct spatial scales:

**1. Data Acquisition and Indicator Processing**  
First, we established the thermal and physical baseline of the cities. Using Google Earth Engine, we pulled summer 2025 data from Landsat 8 (30m resolution) to calculate two key metrics:

*   **Vegetation (NDVI):** To measure the density of green infrastructure.
*   **Land Surface Temperature (LST):** To map the thermal landscape. To ensure accuracy across different satellite capture dates, we normalized the LST data by calculating the deviation from the spatial mean of each image, giving us a clean look at localized heat anomalies rather than day-to-day weather spikes. 

For urban morphology, we integrated the RUBCLIM Local Climate Zone (LCZ) dataset (100m resolution), which categorizes the physical build of the city into classes ranging from "Compact High-Rise" to "Sparsely Built."

**2. Pedestrian Accessibility (PST Engine)**  
To figure out how human-friendly these environments actually are, we couldn't just look at satellite pixels; we had to simulate how people move. Using OSM walking networks, we built a Python-based graph network to calculate Pedestrian Space Syntax (PST) accessibility. 

We identified "cooling hubs" (neighborhoods with an NDVI above the city median) and mapped their center points. The algorithm then simulated human navigation, counting the number of topological turns required to reach these cooling zones from any point on the street network. Paths requiring fewer than three turns received higher accessibility scores, mathematically penalizing areas that are cut off from blue-green infrastructure. 

**3. Macro-Scale Clustering (Neighborhoods)**  
With our variables set (Mean LST, Mean NDVI, LCZ class, and PST score), we aggregated the data up to the neighborhood level using zonal statistics. Because our data was mixed (blending continuous numbers like temperature with categorical tags like LCZ codes), standard clustering methods wouldn't work. 

Instead, we ran the data through R using a Gower distance matrix, which is specifically designed to handle mixed data types. We then applied Partitioning Around Medoids (PAM) clustering to group the neighborhoods into four distinct microclimate profiles.

**4. Micro-Scale Clustering (Street Segments)**  
Finally, we brought the analysis down to the human scale to find the most vulnerable individual streets. Using Python and Geopandas, we slightly buffered the OSM street linestrings and extracted the exact LST and NDVI pixel values sitting directly under the pavement, while sampling the LCZ classification at the start point of each street. 

After dealing with missing values and standardizing the numeric data, we used an elbow plot to determine the optimal number of clusters and ran a MiniBatch K-Means clustering algorithm. To verify that our math matched physical reality, the pipeline randomly sampled 50 streets from our most extreme clusters and exported their coordinates to a CSV, allowing us to drop into Google Street View and qualitatively inspect the physical characteristics of the hottest streets in the city.

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

# Read the layer from the GeoPackage
invisible(capture.output(
  roads_data <- st_read("data/rotterdam/rotterdam_roads_clustered.gpkg",
                        layer = "rotterdam_roads_data")
))

# Randomly select 500 rows
set.seed(42)  # remove or change this if you don't want reproducible sampling
roads_sample <- roads_data %>% slice_sample(n = 500)

# Colors for clusters 0-5
cluster_colors <- c(
  "0" = "#4C72B0",
  "1" = "#65C3A1",
  "2" = "#FC9964",
  "3" = "#869DC5",
  "4" = "#E597C0",
  "5" = "#D9C04C"
)

cluster3d_scaled <- plot_ly(
  x = roads_sample$lst,
  y = roads_sample$ndvi,
  z = roads_sample$Bww1kM,
  type = "scatter3d",
  mode = "markers",
  color = as.factor(roads_sample$cluster),
  colors = cluster_colors,
  marker = list(size = 3)
) %>%
  layout(
    title = "Street clusters in Rotterdam",
    scene = list(
      xaxis = list(title = "LST"),
      yaxis = list(title = "NDVI"),
      zaxis = list(title = "Betweenness")
    )
  )
cluster3d_scaled
```


## Small-scale output: Rotterdam

After completing step 1 and 2, the following results were obtained, which formed the basis for further analysis.

```{=html}
{{< include components/rotterdam-maps.html >}}
```


From the analysis, three neighbourhoods stand out as showing extreme variation relative to the rest of the city: **Bloemhof, Afrikaanderwijk, and Hillesluis**.
```{=html}
{{< include components/rotterdam-cluster-maps.html >}}
```






## Small-scale output: Guangzhou

```{=html}
{{< include components/guangzhou-maps.html >}}
```


```{=html}
{{< include components/yuexiu-district-china.html >}}
```


# Results

## Rotterdam: Street-view and Nature based solutions

```{=html}
{{< include components/street-view-rotterdam.html >}}
```

## Gunagzhou: Street-view and Nature based solutions

```{=html}
{{< include components/yuexiu-street-view.html >}}
```


```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

# Read the layer from the GeoPackage
invisible(capture.output(
  gzhou_roads_data <- st_read("data/guangzhou/guangzhou_roads_clustered.gpkg",
                        layer = "guangzhou_roads_data")
))

# Randomly select 500 rows
set.seed(42)  # remove or change this if you don't want reproducible sampling
gzhou_roads_sample <- gzhou_roads_data %>% slice_sample(n = 500)

# Colors for clusters 0-5
cluster_colors <- c(
  "0" = "#4C72B0",
  "1" = "#65C3A1",
  "2" = "#FC9964",
  "3" = "#869DC5",
  "4" = "#E597C0",
  "5" = "#D9C04C"
)

cluster3d_scaled <- plot_ly(
  x = gzhou_roads_sample$lst,
  y = gzhou_roads_sample$ndvi,
  z = gzhou_roads_sample$Bww1kM,
  type = "scatter3d",
  mode = "markers",
  color = as.factor(gzhou_roads_sample$cluster),
  colors = cluster_colors,
  marker = list(size = 3)
) %>%
  layout(
    title = "Street clusters in Guangzhou",
    scene = list(
      xaxis = list(title = "LST"),
      yaxis = list(title = "NDVI"),
      zaxis = list(title = "Betweenness")
    )
  )
cluster3d_scaled
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

# Convert a hex color to an rgba() string with a given alpha (reused from earlier)
hex_to_rgba <- function(hex, alpha) {
  rgb_vals <- col2rgb(hex)
  sprintf("rgba(%d,%d,%d,%.2f)", rgb_vals[1], rgb_vals[2], rgb_vals[3], alpha)
}

# Violin plots for LST and NDVI, one subplot per variable, one violin per city within each
# data_list: named list, e.g. list(Rotterdam = roads_data, Guangzhou = gzhou_roads_data)
plot_lst_ndvi_violin <- function(data_list, colors = c("#4C72B0", "#FC9964"),
                                  line_opacity = 0.8, fill_opacity = 0.4,
                                  title = "Distribution of LST and NDVI by city") {

  combined <- bind_rows(lapply(names(data_list), function(city_name) {
    data_list[[city_name]] %>%
      st_drop_geometry() %>%
      select(lst, ndvi) %>%
      mutate(city = city_name)
  })) %>%
    mutate(city = factor(city, levels = names(data_list)))  # fixed left-to-right order

  # Build one violin subplot for a given variable (e.g. "lst" or "ndvi")
  build_panel <- function(var, y_title, show_legend) {
    p <- plot_ly()
    for (i in seq_along(data_list)) {
      city_name <- names(data_list)[i]
      sub <- combined %>% filter(city == city_name)

      p <- p %>% add_trace(
        type = "violin",
        x = sub$city,
        y = sub[[var]],
        name = city_name,
        legendgroup = city_name,
        scalegroup = var,        # violins within a panel share scaling
        box = list(visible = TRUE),
        meanline = list(visible = TRUE),
        line = list(color = hex_to_rgba(colors[i], line_opacity)),
        fillcolor = hex_to_rgba(colors[i], fill_opacity),
        points = FALSE,
        showlegend = show_legend
      )
    }
    p %>% layout(
      yaxis = list(title = y_title),
      xaxis = list(title = "")
    )
  }

  p_lst  <- build_panel("lst",  "LST",  show_legend = TRUE)
  p_ndvi <- build_panel("ndvi", "NDVI", show_legend = FALSE)

  subplot(p_lst, p_ndvi, nrows = 1, titleY = TRUE, margin = 0.06) %>%
    layout(
      title = title,
      violinmode = "group",
      legend = list(title = list(text = "City"))
    )
}

plot_lst_ndvi_violin(list(Rotterdam = roads_data, Guangzhou = gzhou_roads_data))
```

```{r message=FALSE, warning=FALSE}
# LST/NDVI correlation for Gzhou

library(plotly)

# Correlation coefficient
correlation <- cor(gzhou_roads_sample$lst, gzhou_roads_sample$ndvi, use = "complete.obs")

# Linear model for the trend line
lm_fit <- lm(ndvi ~ lst, data = gzhou_roads_sample)
lst_range <- seq(min(gzhou_roads_sample$lst, na.rm = TRUE),
                  max(gzhou_roads_sample$lst, na.rm = TRUE),
                  length.out = 100)
trend_pred <- predict(lm_fit, newdata = data.frame(lst = lst_range))

scatter_lst_ndvi <- plot_ly(
  data = gzhou_roads_sample,
  x = ~lst,
  y = ~ndvi,
  type = "scatter",
  mode = "markers",
  color = ~as.factor(cluster),
  colors = cluster_colors,
  marker = list(size = 6, opacity = 0.7),
  name = ~as.factor(cluster)
) %>%
  add_trace(
    x = lst_range,
    y = trend_pred,
    type = "scatter",
    mode = "lines",
    line = list(color = "#000000", width = 2, dash = "dash"),
    name = "Trend line",
    inherit = FALSE   # <-- prevents it from inheriting the marker settings above
  ) %>%
  layout(
    title = paste0("Guangzhou: LST vs NDVI (r = ", round(correlation, 3), ")"),
    xaxis = list(title = "LST"),
    yaxis = list(title = "NDVI")
  )

scatter_lst_ndvi
```

```{r message=FALSE, warning=FALSE}
# LST/NDVI correlation for Rdam
library(plotly)

# Correlation coefficient
correlation <- cor(roads_sample$lst, roads_sample$ndvi, use = "complete.obs")

# Linear model for the trend line
lm_fit <- lm(ndvi ~ lst, data = roads_sample)
lst_range <- seq(min(roads_sample$lst, na.rm = TRUE),
                  max(roads_sample$lst, na.rm = TRUE),
                  length.out = 100)
trend_pred <- predict(lm_fit, newdata = data.frame(lst = lst_range))

scatter_lst_ndvi <- plot_ly(
  data = roads_sample,
  x = ~lst,
  y = ~ndvi,
  type = "scatter",
  mode = "markers",
  color = ~as.factor(cluster),
  colors = cluster_colors,
  marker = list(size = 6, opacity = 0.7),
  name = ~as.factor(cluster)
) %>%
  add_trace(
    x = lst_range,
    y = trend_pred,
    type = "scatter",
    mode = "lines",
    line = list(color = "#000000", width = 2, dash = "dash"),
    name = "Trend line",
    inherit = FALSE   # <-- prevents it from inheriting the marker settings above
  ) %>%
  layout(
    title = paste0("Rotterdam: LST vs NDVI (r = ", round(correlation, 3), ")"),
    xaxis = list(title = "LST"),
    yaxis = list(title = "NDVI")
  )

scatter_lst_ndvi
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

lcz_labels <- c(
  "0"  = "No data",
  "1"  = "Compact highrise",
  "2"  = "Compact midrise",
  "3"  = "Compact lowrise",
  "4"  = "Open highrise",
  "5"  = "Open midrise",
  "6"  = "Open lowrise",
  "7"  = "Lightweight lowrise",
  "8"  = "Large lowrise",
  "9"  = "Sparsely built",
  "10" = "Heavy industry",
  "11" = "Dense trees (LCZ A)",
  "12" = "Scattered trees (LCZ B)",
  "13" = "Bush, scrub (LCZ C)",
  "14" = "Low plants (LCZ D)",
  "15" = "Bare rock/paved (LCZ E)",
  "16" = "Bare soil/sand (LCZ F)",
  "17" = "Water (LCZ G)"
)
lcz_short_labels <- c("0" = "no data", setNames(paste("LCZ", 1:17), as.character(1:17)))

# % of a city's TOTAL road length per LCZ (no cluster grouping)
city_lcz_pct <- function(roads_data, all_lcz = names(lcz_labels)) {
  roads_data$seg_length <- as.numeric(st_length(roads_data))

  df <- roads_data %>%
    st_drop_geometry() %>%
    mutate(lcz = ifelse(is.na(lcz), "0", as.character(lcz))) %>%
    group_by(lcz) %>%
    summarise(total_length = sum(seg_length, na.rm = TRUE), .groups = "drop") %>%
    mutate(pct = 100 * total_length / sum(total_length))

  tibble(lcz = all_lcz) %>%
    left_join(df, by = "lcz") %>%
    mutate(
      pct = ifelse(is.na(pct), 0, pct),
      lcz_full = lcz_labels[lcz],
      lcz_short = lcz_short_labels[lcz]
    ) %>%
    arrange(match(lcz, all_lcz))
}

# Convert a hex color (e.g. "#4C72B0") to an rgba() string with a given alpha
hex_to_rgba <- function(hex, alpha) {
  rgb_vals <- col2rgb(hex)
  sprintf("rgba(%d,%d,%d,%.2f)", rgb_vals[1], rgb_vals[2], rgb_vals[3], alpha)
}

plot_city_lcz_radar <- function(data_list, colors = c("#4C72B0", "#FC9964"),
                                 line_opacity = 0.6, fill_opacity = 0.1,
                                 title = "LCZ composition of total road network length",
                                 title_pad = 16) {
  fig <- plot_ly(type = 'scatterpolar', fill = 'toself')

  for (i in seq_along(data_list)) {
    city_name <- names(data_list)[i]
    pct_df <- city_lcz_pct(data_list[[i]])

    theta_vals <- c(pct_df$lcz_short, pct_df$lcz_short[1])
    r_vals <- c(pct_df$pct, pct_df$pct[1])
    hover_labels <- c(pct_df$lcz_full, pct_df$lcz_full[1])

    fig <- fig %>%
      add_trace(
        r = r_vals,
        theta = theta_vals,
        name = city_name,
        text = hover_labels,
        hovertemplate = paste0("<b>", city_name, "</b><br>%{text}: %{r:.1f}%<extra></extra>"),
        line = list(color = hex_to_rgba(colors[i], line_opacity)),
        marker = list(color = hex_to_rgba(colors[i], line_opacity)),
        fillcolor = hex_to_rgba(colors[i], fill_opacity)
        # no trace-level `opacity` — line/marker and fill now control their own alpha independently
      )
  }

  fig %>%
    layout(
      polar = list(
        domain = list(y = c(0, 0.88)),
        radialaxis = list(visible = TRUE, ticksuffix = "%")
      ),
      title = list(text = title, y = 0.97, yanchor = "top"),
      margin = list(t = title_pad)
    )
}

plot_city_lcz_radar(list(Rotterdam = roads_data, Guangzhou = gzhou_roads_data))
```

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(plotly)

lcz_labels <- c(
  "0"  = "No data",
  "1"  = "Compact highrise",
  "2"  = "Compact midrise",
  "3"  = "Compact lowrise",
  "4"  = "Open highrise",
  "5"  = "Open midrise",
  "6"  = "Open lowrise",
  "7"  = "Lightweight lowrise",
  "8"  = "Large lowrise",
  "9"  = "Sparsely built",
  "10" = "Heavy industry",
  "11" = "Dense trees (LCZ A)",
  "12" = "Scattered trees (LCZ B)",
  "13" = "Bush, scrub (LCZ C)",
  "14" = "Low plants (LCZ D)",
  "15" = "Bare rock/paved (LCZ E)",
  "16" = "Bare soil/sand (LCZ F)",
  "17" = "Water (LCZ G)"
)

lcz_short_labels <- c("0" = "no data", setNames(paste("LCZ", 1:17), as.character(1:17)))

cluster_lcz_pct <- function(roads_data, all_lcz = names(lcz_labels)) {
  roads_data$seg_length <- as.numeric(st_length(roads_data))

  df <- roads_data %>%
    st_drop_geometry() %>%
    mutate(
      cluster = as.character(cluster),
      lcz = ifelse(is.na(lcz), "0", as.character(lcz))
    ) %>%
    group_by(cluster, lcz) %>%
    summarise(total_length = sum(seg_length, na.rm = TRUE), .groups = "drop") %>%
    group_by(cluster) %>%
    mutate(pct = 100 * total_length / sum(total_length)) %>%
    ungroup()

  clusters <- sort(unique(df$cluster))
  full_grid <- expand.grid(cluster = clusters, lcz = all_lcz, stringsAsFactors = FALSE)

  full_grid %>%
    left_join(df, by = c("cluster", "lcz")) %>%
    mutate(
      pct = ifelse(is.na(pct), 0, pct),
      lcz_full = lcz_labels[lcz],
      lcz_short = lcz_short_labels[lcz]
    ) %>%
    arrange(cluster, match(lcz, all_lcz))
}

# Faceted radar with explicit pixel sizing so panels stay square and fill their allotted space
plot_cluster_lcz_radar_facets <- function(roads_data, colors = cluster_colors,
                                           ncols = 2,
                                           panel_size = 280,   # px — each polar panel is panel_size x panel_size
                                           title_band = 40,    # px reserved above each panel for its title
                                           gap = 100,            # px gap between panels
                                           main_title_margin = 60,  # px reserved at top for the overall title
                                           side_margin = 48,
                                           bottom_margin = 20,
                                           title = "LCZ composition by cluster") {
  pct_df <- cluster_lcz_pct(roads_data)
  clusters <- sort(unique(pct_df$cluster))
  n <- length(clusters)
  nrows <- ceiling(n / ncols)

  r_max <- ceiling(max(pct_df$pct, na.rm = TRUE) / 5) * 5

  # Plot-area (inside margins) dimensions, built up from panel_size so squares stay square
  plot_area_width  <- ncols * panel_size + (ncols - 1) * gap
  plot_area_height <- nrows * (panel_size + title_band) + (nrows - 1) * gap

  fig_width  <- plot_area_width + 2 * side_margin
  fig_height <- plot_area_height + main_title_margin + bottom_margin

  fig <- plot_ly(type = 'scatterpolar', fill = 'toself')
  polar_layouts <- list()
  annotations <- list()

  for (i in seq_along(clusters)) {
    cl <- clusters[i]
    row_idx <- (i - 1) %/% ncols
    col_idx <- (i - 1) %% ncols

    x0_px <- col_idx * (panel_size + gap)
    x1_px <- x0_px + panel_size

    block_top_px    <- row_idx * (panel_size + title_band + gap)
    polar_top_px    <- block_top_px + title_band
    polar_bottom_px <- polar_top_px + panel_size

    # Fractions are relative to the plot area (same space "paper" coords use), not the full figure
    x0 <- x0_px / plot_area_width
    x1 <- x1_px / plot_area_width
    y1 <- 1 - polar_top_px / plot_area_height
    y0 <- 1 - polar_bottom_px / plot_area_height
    title_y <- 1 - block_top_px / plot_area_height

    polar_name <- if (i == 1) "polar" else paste0("polar", i)

    sub <- pct_df %>% filter(cluster == cl)
    theta_vals <- c(sub$lcz_short, sub$lcz_short[1])
    r_vals <- c(sub$pct, sub$pct[1])
    hover_labels <- c(sub$lcz_full, sub$lcz_full[1])

    fig <- fig %>%
      add_trace(
        r = r_vals,
        theta = theta_vals,
        subplot = polar_name,
        name = paste("Cluster", cl),
        text = hover_labels,
        hovertemplate = paste0("<b>Cluster ", cl, "</b><br>%{text}: %{r:.1f}%<extra></extra>"),
        line = list(color = colors[[cl]]),
        marker = list(color = colors[[cl]]),
        opacity = 0.7,
        showlegend = FALSE
      )

    polar_layouts[[polar_name]] <- list(
      domain = list(x = c(x0, x1), y = c(y0, y1)),
      radialaxis = list(visible = TRUE, range = c(0, r_max), ticksuffix = "%", tickfont = list(size = 8)),
      angularaxis = list(tickfont = list(size = 8))
    )

    annotations[[i]] <- list(
      text = paste("Cluster", cl),
      x = (x0 + x1) / 2, y = title_y,
      xanchor = "center", yanchor = "top",
      xref = "paper", yref = "paper",
      showarrow = FALSE,
      font = list(size = 13)
    )
  }

  layout_args <- c(
    list(fig),
    polar_layouts,
    list(
      title = title,
      annotations = annotations,
      width = fig_width,
      height = fig_height,
      margin = list(t = main_title_margin, b = bottom_margin, l = side_margin, r = side_margin)
    )
  )

  do.call(layout, layout_args)
}
```

```{r message=FALSE, warning=FALSE}
plot_cluster_lcz_radar_facets(roads_data, title = "Rotterdam: LCZ composition by cluster")
```

```{r message=FALSE, warning=FALSE}
plot_cluster_lcz_radar_facets(gzhou_roads_data, title = "Guangzhou: LCZ composition by cluster")
```

# Discussion

# Conclusion
